### 2023-02-14 source/intro.Rnw
paper VI revision and proofs

\startPaperSummary{PaperVI}{%
  Pitfalls and Potentials in Simulation Studies}{%
  Samuel Pawel, Lucas Kook, Kelly Reeve \\
  \textit{arXiv preprint}, 2022.
  doi:\href{https://doi.org/10.48550/arXiv.2203.13076}{10.48550/arXiv.2203.13076}.}

@@

\phantomsection
\addcontentsline{toc}{section}{Data and software}
\section*{Data and software}
The CC-By 4.0 licensed data from the ``Reproducibility Project: Cancer Biology''
\citep{Errington2021} were downloaded from
\url{https://doi.org/10.17605/osf.io/e5nvr}. The relevant variables were then
extracted from the file ``\texttt{RP\_CB Final Analysis - Effect level
  data.csv}''. All analyses from this chapter were conducted in the R
programming language version \Sexpr{paste(version$major, version$minor, sep =
  ".")} \citep{R}. The packages \texttt{dplyr} \citep{Wickham2022},
\texttt{ggplot2} \citep{Wickham2016}, \texttt{gridExtra} \citep{Auguie2017},
\texttt{knitr} \citep{Xie2022}, \texttt{ReplicationSuccess} \citep{Held2020},
\texttt{scales} \citep{Wickham2022b}, \texttt{UpSetR} \citep{Gehlenborg2019},
and \texttt{xtable} \citep{Dahl2019} were used. Code and data to reproduce this
thesis are available at \url{https://github.com/SamCH93/thesis}. A snapshot of
the git repository is archived at \url{https://doi.org/10.5281/zenodo.XXXXXXX}.

>>>>

\startPaperSummary{PaperVI}{%
  Pitfalls and Potentials in Simulation Studies}{%
  Samuel Pawel, Lucas Kook, Kelly Reeve \\
  \textit{Biometrical Journal}, 2023.
  doi:\href{https://doi.org/10.1002/bimj.202200091}{10.1002/bimj.202200091}.}

@@

\phantomsection
\addcontentsline{toc}{section}{Data and software}
\section*{Data and software}
The CC-By 4.0 licensed data from the ``Reproducibility Project: Cancer Biology''
\citep{Errington2021} were downloaded from
\url{https://doi.org/10.17605/osf.io/e5nvr}. The relevant variables were then
extracted from the file ``\texttt{RP\_CB Final Analysis - Effect level
  data.csv}''. All analyses from this chapter were conducted in the R
programming language version \Sexpr{paste(version$major, version$minor, sep =
  ".")} \citep{R}. The packages \texttt{dplyr} \citep{Wickham2022},
\texttt{ggplot2} \citep{Wickham2016}, \texttt{gridExtra} \citep{Auguie2017},
\texttt{knitr} \citep{Xie2022}, \texttt{ReplicationSuccess} \citep{Held2020},
\texttt{scales} \citep{Wickham2022b}, \texttt{UpSetR} \citep{Gehlenborg2019},
and \texttt{xtable} \citep{Dahl2019} were used. Code and data to reproduce this
thesis are available at \url{https://github.com/SamCH93/thesis}. A snapshot of
the git repository is archived at \url{https://doi.org/10.5281/zenodo.7464296}.

### 2023-02-14 source/paper6.tex
reorganizatin for reproducibility
paper VI revision and proofs

This article is structured as follows: We first give an illustrative list of
QRPs related to comparative simulation studies (Section \ref{sec6:QRP}). With an
exemplary simulation study, we then show h?ow easy it is to present a novel,
made-up method as an improvement over others if QRPs are employed and \emph{a
  priori} simulation plans remain undisclosed (Section \ref{sec6:study}). The
main inspiration for this work is drawn from similar illustrative studies which
have been conducted by \citet{Yousefi2009} and \citet{Jelizarow2010} for
benchmarking studies, and by \citet{Simmons2011} in the context of $p$-hacking
in psychological research. Recently, \citet{Niessl2021} and \citet{Ullmann2022}
expanded on QRPs in benchmarking studies with the latter also including
simulation studies. In Section~\ref{sec6:recommendations}, we then provide
concrete suggestions for researchers, reviewers, editors and funding bodies to
alleviate the issues of QRPs and improve the methodological quality of
comparative simulation studies. Section~\ref{sec6:discussion} closes with
limitations and concluding remarks.

@@

\begin{table}[!htb]
  \caption{Types of questionable research practices (QRPs) in comparative
    simulation studies at different stages of the research process. A QRP
    becomes more problematic if combined with a related QRP, especially a
    reporting QRP.}
  \label{table6:QRPs}
  \centering
  \begin{tabular}{p{.05\textwidth} p{.11\textwidth} p{.75\textwidth}}
    \toprule
    \textbf{Tag} & \textbf{Related} & \textbf{Type of QRP} \\
    \midrule
    \multicolumn{2}{p{.15\textwidth}}{\textit{Design}} & \\
    D1 & E1, R1 & Not/vaguely defining objectives of simulation study \\
    D2 & E2, R1 & Not/vaguely defining data-generating process \\
    D3 & E3, E4, R1 & Not/vaguely defining which methods will be compared and how their
                      parameters are specified \\
    D4 & E1, E5, R1 & Not/vaguely defining estimands of interest \\
    D5 & E1, E5, R1 & Not/vaguely defining evaluation criteria \\
    D6 & E6, R1 & Not/vaguely defining how to handle missing values
                  (for example, due to non-convergence of methods) \\
    D7 & E7, E8, R3 & Not justifying number of simulations \\[1em]

@@

\section*{Acknowledgements}
We thank Eva Furrer, Malgorzata Roos and Torsten Hothorn for helpful discussion
and comments on the simulation protocol and drafts of the manuscript. We also
thank the anonymous referees and the associate editor for constructive and
valuable comments that improved the manuscript substantially. Our
acknowledgement of these individuals does not imply their endorsement of this
article. The authors declare that they do not have any conflicts of interest.

@@

  \item \textbf{AUC.} The AUC is given by
  $$\text{AUC} = \max\{\text{PI}, 1 - \text{PI}\}$$
        with $$\text{PI} = \widehat\Prob(Y_i \geq Y_j \given \rx_i, \rx_j), \; i,j = 1, \dots, n_{\text{test}},$$
        where $Y_i$ and $Y_j$ denote case and non-case, respectively. The AUC is
        related to the area under the receiver-operating-characteristic (ROC)
        curve \citep{Steyerberg2019}. It will be used as a measure of
        discrimination and values closer to one indicate better discriminative
        ability. Discrimination describes the ability of a prediction model to
        discriminate between cases and non-cases. Other discrimination measures,
        such as accuracy, sensitivity, specificity, etc., are not considered
        because we want to evaluate predictive performance in terms of
        probabilistic predictions instead of point predictions/classification.

@@

The penalized log-likelihood for \ainet{} for a single observation $(\ry, \rx)$
is defined as
$$\ell_{\text{AINET}}(\beta_0, \shiftparm; \ry, \rx, \alpha, \lambda, \wvec) =
  \ell(\beta_0, \shiftparm; \ry, \rx) + \aipen$$
where
$$\ell(\beta_0, \shiftparm; \ry, \rx) =
\ry \log\left(\expit\left\{\beta_0 + \linpred\right\}\right) + (1 - \ry) \log\left(1 - \expit\left\{\beta_0 + \linpred\right\}\right)$$
denotes the log-likelihood of a binomial GLM and $\wvec$ is derived from a
random forest variable importance measure $\IMP$ as
$$w_j = 1 - \left(\frac{\IMP_j}{\sum_{k=1}^p \IMP_k}\right)^\gamma,$$
where we transform $\IMP$ to be non-negative via
$$\IMP_j = \max\{0, \widetilde\IMP_j\}$$
and $\gamma$ is a hyperparameter for the influence of the weights similar to
$\gamma$ hyperparameter of the adaptive elastic net. \ainet{} is fitted by
maximizing its penalized log-likelihood assuming i.i.d. observations
$\{(\ry_i, \rx_i)\}_{i=1}^n$, \ie
$$\arg\max_{\beta_0, \shiftparm} \sum_{i = 1}^n \ell_{\text{AINET}}(\beta_0,
\shiftparm; \ry_i, \rx_i, \alpha, \lambda, \wvec).$$

@@

Figure~\ref{fig6:tienll} shows the differences in area under the curve (AUC)
between \ainet{} and the other methods stratified by simulation conditions. As
with the other estimands, \ainet{} shows virtually identical performance as EN
regression across all simulation conditions. \ainet{} seems to outperform RF
across most simulation conditions, with the exception of a conditions with low
sample size ($n = 100$), medium prevalence ($\mbox{prev} = 0.05$), and low
events per variable ($\mbox{EPV} \leq 1$). GLM, typically outperforms \ainet{}
conditions with small to medium sample size ($n \leq 500$), and also in
conditions with larger sample size when the events per variable is normal to
high ($\mbox{EPV} \geq 10$) and the prevalence is small ($\mbox{prev} = 0.01$).
Finally, the AEN is worse with respect to AUC than \ainet{} across all
simulation conditions.

@@

\begin{landscape}
\begin{figure}[!ht]
\center
\includegraphics[width=0.9\linewidth]{images/paper6/tie-fighter_brier_sparsity0.pdf}
\caption{Tie-fighter plot for the difference in Brier score between any method
  on the $y$-axis and \ainet{}. The 95\% confidence intervals are adjusted per
  simulation condition using the single-step method. Lower values indicate
  better performance of \ainet{}. } \label{fig6:tiebrier}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}[!ht]
\center
\includegraphics[width=0.9\linewidth]{images/paper6/tie-fighter_scaledBrier_sparsity0.pdf}
\caption{Tie-fighter plot for the difference in scaled Brier score between any
  method on the $y$-axis and \ainet{}. The 95\% confidence intervals are adjusted
  per simulation condition using the single-step method. Larger
  values indicate better performance of \ainet{}. } \label{fig6:tiesbrier}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}[!ht]
\center
\includegraphics[width=0.9\linewidth]{images/paper6/tie-fighter_nll_sparsity0.pdf}
\caption{Tie-fighter plot for the difference in log-score between any method on
  the $y$-axis and \ainet{}. The 95\% confidence intervals are adjusted per
  simulation condition using the single-step method. Lower values indicate
  better performance of \ainet{}. } \label{fig6:tienll}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}[!ht]
\center
\includegraphics[width=0.9\linewidth]{images/paper6/calibration-cslope_sparsity0.pdf}
\caption{Boxplots of calibration slopes stratified by method and simulation
  conditions. Mean calibration slope is indicated by a cross. A value of one
  indicates optimal calibration. Percentage of simulations where calibration
  slope could not be estimated (due to extreme predictions or complete
  separation) are also indicated.} \label{fig6:cslope}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}[!ht]
\center
\includegraphics[width=0.9\linewidth]{images/paper6/calibration-clarge_sparsity0.pdf}
\caption{Boxplots of calibration in the large stratified by method and
  simulation conditions. Mean calibration in the large is indicated by a cross.
  A value of zero indicates optimal calibration in the large. Percentage of
  simulations where calibration in the large could not be estimated (due to
  extreme predictions or complete separation) are also
  indicated.} \label{fig6:clarge}
\end{figure}
\end{landscape}

>>>>

This article is structured as follows: We first give an illustrative list of
QRPs related to comparative simulation studies (Section \ref{sec6:QRP}). With an
exemplary simulation study, we then show how easy it is to present a novel,
made-up method as an improvement over others if QRPs are employed and \emph{a
  priori} simulation plans remain undisclosed (Section \ref{sec6:study}). The
main inspiration for this work is drawn from similar illustrative studies which
have been conducted by \citet{Yousefi2009} and \citet{Jelizarow2010} for
benchmarking studies, and by \citet{Simmons2011} in the context of $p$-hacking
in psychological research. Recently, \citet{Niessl2021} and \citet{Ullmann2022}
expanded on QRPs in benchmarking studies with the latter also including
simulation studies. In Section~\ref{sec6:recommendations}, we then provide
concrete suggestions for researchers, reviewers, editors and funding bodies to
alleviate the issues of QRPs and improve the methodological quality of
comparative simulation studies. Section~\ref{sec6:discussion} closes with
limitations and concluding remarks.

@@

\begin{table}[!htb]
  \caption{Types of questionable research practices (QRPs) in comparative
    simulation studies at different stages of the research process. A QRP
    becomes more problematic if combined with a related QRP, especially a
    reporting QRP.}
  \label{table6:QRPs}
  \centering
  \begin{tabular}{p{.05\textwidth} p{.115\textwidth} p{.75\textwidth}}
    \toprule
    \textbf{Tag} & \textbf{Related} & \textbf{Type of QRP} \\
    \midrule
    \multicolumn{2}{p{.15\textwidth}}{\textit{Design}} & \\
    D1 & E1, R1 & Not/vaguely defining objectives of simulation study \\
    D2 & E2, R1 & Not/vaguely defining data-generating process \\
    D3 & E3, E4, R1 & Not/vaguely defining which methods will be compared and how their
                      parameters are specified \\
    D4 & E1, E5, R1 & Not/vaguely defining estimands of interest \\
    D5 & E1, E5, R1 & Not/vaguely defining evaluation criteria \\
    D6 & E6, R1 & Not/vaguely defining how to handle missing values
                  (for example, due to non-convergence of methods) \\
    D7 & E7, E8, R3 & Not justifying number of simulations \\[1em]

@@

\section*{Acknowledgements}
We thank Eva Furrer, Malgorzata Roos and Torsten Hothorn for helpful discussion
and comments on the simulation protocol and drafts of the manuscript. We also
thank the anonymous referees and the associate editor for constructive and
valuable comments that improved the manuscript substantially. Our
acknowledgement of these individuals does not imply their endorsement of this
article.

@@

  \item \textbf{AUC.} The AUC is the area under the
        receiver-operating-characteristic (ROC) curve \citep{Steyerberg2019}. It
        will be used as a measure of discrimination and values closer to one
        indicate better discriminative ability. Discrimination describes the
        ability of a prediction model to discriminate between cases and
        non-cases. Other discrimination measures, such as accuracy, sensitivity,
        specificity, etc., are not considered because we want to evaluate
        predictive performance in terms of probabilistic predictions instead of
        point predictions/classification.

@@

The penalized log-likelihood for \ainet{} for a single observation $(\ry, \rx)$
is defined as
$$\ell_{\text{AINET}}(\beta_0, \shiftparm; \ry, \rx, \alpha, \lambda, \wvec) =
  \ell(\beta_0, \shiftparm; \ry, \rx) + \aipen$$
where
$$\ell(\beta_0, \shiftparm; \ry, \rx) =
\ry \log\left(\expit\left\{\beta_0 + \linpred\right\}\right) + (1 - \ry) \log\left(1 - \expit\left\{\beta_0 + \linpred\right\}\right)$$
denotes the log-likelihood of a binomial GLM and $\wvec$ is derived from a
random forest variable importance measure $\widetile{IMP}$ as
$$w_j = 1 - \left(\frac{\IMP_j}{\sum_{k=1}^p \IMP_k}\right)^\gamma,$$
where we transform $\IMP$ to be non-negative via
$$\IMP_j = \max\{0, \widetilde\IMP_j\}$$
and $\gamma$ is a hyperparameter for the influence of the weights similar to
$\gamma$ hyperparameter of the adaptive elastic net. \ainet{} is fitted by
maximizing its penalized log-likelihood assuming i.i.d. observations
$\{(\ry_i, \rx_i)\}_{i=1}^n$, \ie
$$\arg\max_{\beta_0, \shiftparm} \sum_{i = 1}^n \ell_{\text{AINET}}(\beta_0,
\shiftparm; \ry_i, \rx_i, \alpha, \lambda, \wvec).$$

@@

Figure~\ref{fig6:tieAUC} shows the differences in area under the curve (AUC)
between \ainet{} and the other methods stratified by simulation conditions. As
with the other estimands, \ainet{} shows virtually identical performance as EN
regression across all simulation conditions. \ainet{} seems to outperform RF
across most simulation conditions, with the exception of a conditions with low
sample size ($n = 100$), medium prevalence ($\mbox{prev} = 0.05$), and low
events per variable ($\mbox{EPV} \leq 1$). GLM, typically outperforms \ainet{}
conditions with small to medium sample size ($n \leq 500$), and also in
conditions with larger sample size when the events per variable is normal to
high ($\mbox{EPV} \geq 10$) and the prevalence is small ($\mbox{prev} = 0.01$).
Finally, the AEN is worse with respect to AUC than \ainet{} across all
simulation conditions.

@@

\begin{landscape}
\begin{figure}[!ht]
\center
\includegraphics[width=0.9\linewidth]{images/paper6/figureS1.pdf}
\caption{Tie-fighter plot for the difference in Brier score between any method
  on the $y$-axis and \ainet{}. The 95\% confidence intervals are adjusted per
  simulation condition using the single-step method. Lower values indicate
  better performance of \ainet{}. } \label{fig6:tiebrier}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}[!ht]
\center
\includegraphics[width=0.9\linewidth]{images/paper6/figureS2.pdf}
\caption{Tie-fighter plot for the difference in scaled Brier score between any
  method on the $y$-axis and \ainet{}. The 95\% confidence intervals are adjusted
  per simulation condition using the single-step method. Larger
  values indicate better performance of \ainet{}. } \label{fig6:tiesbrier}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}[!ht]
\center
\includegraphics[width=0.9\linewidth]{images/paper6/figureS3.pdf}
\caption{Tie-fighter plot for the difference in log-score between any method on
  the $y$-axis and \ainet{}. The 95\% confidence intervals are adjusted per
  simulation condition using the single-step method. Lower values indicate
  better performance of \ainet{}. } \label{fig6:tienll}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}[!ht]
\center
\includegraphics[width=0.9\linewidth]{images/paper6/figureS4.pdf}
\caption{Tie-fighter plot for the difference in AUC between any method on
  the $y$-axis and \ainet{}. The 95\% confidence intervals are adjusted per
  simulation condition using the single-step method. Larger values indicate
  better performance of \ainet{}. } \label{fig6:tieAUC}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}[!ht]
\center
\includegraphics[width=0.9\linewidth]{images/paper6/figureS5.pdf}
\caption{Boxplots of calibration slopes stratified by method and simulation
  conditions. Mean calibration slope is indicated by a cross. A value of one
  indicates optimal calibration. Percentage of simulations where calibration
  slope could not be estimated (due to extreme predictions or complete
  separation) are also indicated.} \label{fig6:cslope}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}[!ht]
\center
\includegraphics[width=0.9\linewidth]{images/paper6/figureS6.pdf}
\caption{Boxplots of calibration in the large stratified by method and
  simulation conditions. Mean calibration in the large is indicated by a cross.
  A value of zero indicates optimal calibration in the large. Percentage of
  simulations where calibration in the large could not be estimated (due to
  extreme predictions or complete separation) are also
  indicated.} \label{fig6:clarge}
\end{figure}
\end{landscape}
